\cleardoublepage
\thispagestyle{empty}


\vspace*{3cm}

\begin{raggedleft}
    	If it were not for the great variability between individuals, medicine might as well be a science, not an art." \\ \hfill \break
     --- Sir William Oslen\\
\end{raggedleft}

\vspace{4cm}

I am thankful for Andres, Damian, John, and the University of Cambridge's Department of Medicine for hosting me for this master's thesis. I am particularly grateful for Damian for the excellent supervision, for sharing a thorough reviewing mindset and for accompanying me to explore different methods. In Switzerland, I am thankful for Philippe, Martin, and Annie, for their continuous support during my master's at EPFL, through our discussions and the many advice in the projects we undertook together. I am also highly grateful for the CF clinicians and researchers from Hôpital Necker Enfants Malades (France), Arizona Health Sciences Center (US) and Royal Papworth Hospital (UK), for providing an expert feedback on my work. Eventually, I would also like to thank all Project Breathe collaborators for building the richest CF data set of high-frequency physiological measurements in history. I trust that it will help to improve CFers lives condition even further than the current multiple benefits. More personally, as a CF patient and a researcher, it has been hard but also mind-opening to discover the diversity of CFers conditions. Studying CF in the spotlight of the life-changing Trikafta CFTR modulator therapy introduction in Europe was therefore particularly motivating.